### 2021-09-07 input/conclusion.tex
wining space piece by piece
mostly about the implementation

As future work, we plan on extending the logical partitioning protocol
to enable lazy retrieval of partitions. Indeed, using \NAME, every
\process partakes in the propagation of indexes of content and
replicas. We would like to leverage the hierarchical properties of
interconnected autonomous systems to limit the propagation of such
indexes only to interested systems.
%
%% As ongoing and future work, we are currently studying how it might be
%% possible to limit the scope of a partition to one autonomous system at
%% most. Indeed, the gain of propagating changes to \processes belonging to
%% another autonomous system might be negligible as the latency penalty
%% is mainly due to inter autonomous systems' links (\eg trans-continental
%% links).  Exactly, we are investigating how such a change can be
%% formalized and its impact on the current protocol.
%
%autonomous systems. Indeed, \processes could leverage the hierarchical
%properties of the Internet topology~\cite{nur2018geography} to avoid
%flooding the whole systems with control information about all
%partitions.
%\TODO{should we highlight this at the end of the experiment}
%

>>>>

As future work, we plan on extending the logical partitioning protocol
to enable lazy retrieval of partitions \TODO{when a large number of
  objects is involved}. Indeed, using \NAME, every \process partakes
in the propagation of indexes \TODO{which raises scalability
  issues}. We would like to leverage the hierarchical properties of
interconnected autonomous systems to limit the propagation of such
indexes only to interested systems.

%
%% As ongoing and future work, we are currently studying how it might be
%% possible to limit the scope of a partition to one autonomous system at
%% most. Indeed, the gain of propagating changes to \processes belonging to
%% another autonomous system might be negligible as the latency penalty
%% is mainly due to inter autonomous systems' links (\eg trans-continental
%% links).  Exactly, we are investigating how such a change can be
%% formalized and its impact on the current protocol.
%
%autonomous systems. Indeed, \processes could leverage the hierarchical
%properties of the Internet topology~\cite{nur2018geography} to avoid
%flooding the whole systems with control information about all
%partitions.
%\TODO{should we highlight this at the end of the experiment}
%

### 2021-09-07 input/proposal.tex
mostly about the implementation

% \vfill

% \TODO{spacing.}

% \newpage

@@

Algorithm~\ref{algo:ascast} provides the few instructions of \NAME
that implement three-phase purge and best eventual forwarding to
enable dynamic consistent partitioning.

%% Similarly to Algorithm~\ref{algo:add}, local operations \texttt{Add}
%% and \texttt{Del} generate messages that the initiating \process treats
%% identically to receipts of messages from remote \processes.

%% \begin{inparaenum}[(i)]
%% \item \label{algo:most} most of the time, it propagates $\alpha$'s and
%%   $\delta$'s when the partitions allow it;
%% \item \label{algo:sometimes} sometimes, it detects possible
%%   inconsistent partitioning and
%% \item \label{algo:solves}solves it using propagation trees (as opposed
%%   to propagation graphs);
%% \item \label{algo:competition} when required, it triggers competitions
%%   among neighboring partitions.
%% \end{inparaenum}
%% While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
%% stale notifications, (\ref{algo:competition}) tackle the eventual
%% delivery of the best up-to-date partitions.

@@

To implement the removal of stale messages as stated in
Definition~\ref{def:lwo}, each \process maintains a vector of versions
that associates the respective known local counter of each known
source, or has-been source. Each message $m$ carries its source $s$
along with its version $c$ that we subscript $m_{s, c}$.  Each
\process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
% This vector enables the simplest Propagation~I and Propagation~I'.
Its size increases linearly with the number of sources that the
\process has ever known, which is the number of \processes in the
system $\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\Delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such a
$\Delta_{s, c}^{\pi}$ fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such a delivery, messages piggyback monotonically such
increasing control information).

$\delta_{s, c}$ messages used in Propagation~I are more efficient than
$\Delta_{s, c}^\pi$ messages used in Propagation~III, both in
propagation time and traffic. First, the former propagate using every
communication link available in the partition, while the latter
propagate following the dissemination tree of the corresponding
$\alpha$ to purge. Second, the former only piggyback source and
counter while the latter piggyback the path taken by the corresponding
$\alpha$ to purge.  In the worst case, a path includes every \process
in the system but converges to the average diameter of
partitions. Nevertheless, message paths and \NAME synergyze with each
other: Paths tend to be smaller as the number of sources in the system
increases.
%% Bloom filters could introduce another trade-off by further
%% decreasing the size of path data structures at the cost of premature
%% stopped propagation that could impair consistent
%% partitioning~\cite{whitaker2002forwarding}.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale control information, but
also the retrieval of the best up-to-date partitions. In that regard,
both kinds of $\delta$ messages have dual use: since they already
reach the borders of their partition when they remove stale control
information, they also trigger a competition when reaching such
bordering \processes (see Line~\ref{line:ascast_compete}). This simply
consists in sending the current partition through the communication
link from which the \process received the $\delta$. Upon receipt of
this answer, \processes act normally by propagating their changes when
they improve.

\NAME guarantees dynamic consistent partitioning by making extensive
use of \process-to-\process communications. It requires
\begin{inparaenum}[(i)]
\item purging stale control information, hence propagating $\delta$
  messages; and
\item retrieving the closest source, hence propagating $\alpha$
  messages.
\end{inparaenum}
In terms of number of messages, in the average case, a \process $i$
chosen uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.
% \TODO{Multiple receipt and
% multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.

>>>>

@@

Algorithm~\ref{algo:ascast} provides the few instructions of \NAME
that implement three-phase purge and best eventual forwarding to
enable dynamic consistent partitioning. For the sake of simplicity, it
does not handle multiple sessions and associated optimisation such as
data structure sharing and message packing.

@@

As stated in Theorem~\ref{theo:dcp}, every \process must purge stale
messages.  To identify staleness, each \process maintains a vector of
versions that associates the respective known local counter of each
known source, or has-been source. Its size increases linearly with the
number of sources that the \process has ever known, which is the
number of \processes in the system $\mathcal{O}(V)$ in the worst case.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.
Using such a vector, every message $m$ carries its source $s$ along
with its version $c$ that we subscript $m_{s, c}$.  Each \process
delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector of versions constitutes a summary of local deliveries. It
allows a \process to detect inconsistencies in message delivery: the
parent delivered a stale message (see Line~\ref{line:ascast_version}),
or the parent delivered a worse message (see
Line~\ref{line:detectA}). This may happen in stage II of
Figure~\ref{fig:proofB} between $b$ and $c$.

A \process ($c$) detecting an inconsistency disseminates a $\Delta$
message. Such a $\Delta$ must not modify any vector of versions since
the partition of the corresponding $\alpha$ may still exist.
Therefore, it additionally includes and maintains a path $\pi$
initialized to $\alpha$'s one. Such a $\Delta_{s, c}^{\pi}$
transitively tracks and removes $\alpha$ messages that were forwarded
by the detecting \process. The path $\pi$ also enables a quick stop of
looping messages and in turns ensures termination. However, carrying
paths in messages is costly. In the worst case, a message piggybacks
the identity of all \processes in the system. Fortunately, message
paths and \NAME synergyze well with each other, for paths tend to be
smaller as the number of sources in the system increases. Contrarily
to $\alpha$ and $\Delta$ messages, $\delta_{s, c}$ messages do not
carry a path. They
\begin{inparaenum}[(i)]
\item solely rely on versions to operate, and
\item use the whole dissemination graph to propagate in the system, as
  opposed to $\Delta$ that must follow the dissemination tree of the
  corresponding $\alpha$.
\end{inparaenum}

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale control information, but
also the retrieval of the best up-to-date partitions.  In that regard,
both $\delta$ and $\Delta$ messages have dual use: since they already
reach the borders of their partition by design, they also trigger a
competition when reaching such bordering \processes (see
Line~\ref{line:ascast_compete}).
%% This simply consists in sending the
%% current partition through the communication link from which the
%% \process received the $\delta$. Upon receipt of this answer,
%% \processes act normally by propagating their changes when they
%% improve.

%% \NAME guarantees dynamic consistent partitioning by making extensive
%% use of \process-to-\process communications. It requires
%% \begin{inparaenum}[(i)]
%% \item purging stale control information, hence propagating $\delta$
%%   messages; and
%% \item retrieving the closest source, hence propagating $\alpha$
%%   messages.
%% \end{inparaenum}
In terms of number of messages, in the average case, a \process $i$
chosen uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.
% \TODO{Multiple receipt and
% multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.
